\documentclass[12pt]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{array}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{fontspec}
\setmainfont{Noto Sans}

\title{OIP Sentinel: Distributed Database Prototype (Firestore + Aerospike)}
\author{Group Name: Sharpness\\
Members: Faustino Sotomayor, Erika Bicoy, Mark Angelo Dontar, Sanny Fhel Paalisbo, Jomari Sunogan}
\date{November 25, 2025}

\begin{document}
\maketitle

\section{Executive Summary and Project Status}
During Week 1, we integrated Firebase Auth with RBAC via Firestore and implemented step-up 2FA (email OTP) with trusted devices. A Node/Express backend bridges Firestore and Aerospike for dynamic table operations and OTP verification. Core authentication, auditing, and database operations work end-to-end.

\section{System Overview}

\subsection{Features Implemented (or Planned)}
\begin{longtable}{p{0.35\linewidth} p{0.6\linewidth}}
\toprule
\textbf{Feature Name} & \textbf{Status and Progress} \\
\midrule
User Authentication (Sign Up/Login) &
Completed. Email/password and Google Sign-In via Firebase Auth; role/status enforced. \\
RBAC (Admin/Cashier) &
Completed. Firestore \texttt{users/\{uid\}} with \texttt{role} and \texttt{status}; UI gated; role cached safely. \\
2FA (OTP + Trusted Devices) &
Completed. Email OTP via backend; trusted device window under \texttt{users/\{uid\}/trustedDevices}. \\
Audit Logging &
Completed. Admin/cashier logs to Firestore with normalized events and device metadata. \\
Backend DB Manager (REST) &
Completed. Express routes to list/insert/read across Firestore/Aerospike. CORS enabled. \\
Aerospike Integration &
Completed. Namespace \texttt{oipms}; key-based insert; normalized scan reads. \\
Password Reset OTP &
Client-ready; server endpoint planned. \\
ML Object Detection &
Planned; not implemented this week. \\
\bottomrule
\end{longtable}

\subsection{Database Integration}
\textbf{Database Technology:} Firebase Firestore (RBAC, OTP, audit), Aerospike (business data).

\textbf{Schema/Model Progress:}
\begin{itemize}[leftmargin=*]
\item Firestore:
  \begin{itemize}
  \item \texttt{users/\{uid\}}: \texttt{email}, \texttt{role}, \texttt{status}, \texttt{username}[array], \texttt{phone}[array]; subcollection \texttt{trustedDevices/\{deviceId\}}: \texttt{trustedUntil}, \texttt{firstSeenAt}, \texttt{lastSeenAt}.
  \item \texttt{audit\_logs\_admin}, \texttt{audit\_logs\_cashier}: event, timestamp, uid/email/username, route/device details.
  \item \texttt{\_loginOtps/\{uid\}}: \texttt{hash} (SHA-256), \texttt{attempts}, \texttt{maxAttempts}, \texttt{expiresAt}.
  \item \texttt{tables}: metadata for table manager UI.
  \end{itemize}
\item Aerospike:
  \begin{itemize}
  \item Namespace \texttt{oipms}; sets enumerated; records written with app primary key or timestamp fallback.
  \end{itemize}
\end{itemize}

\textbf{Integration Status:}
\begin{itemize}[leftmargin=*]
\item Frontend: Firebase Auth + Firestore reads/writes operational for RBAC, audit, devices.
\item Backend: Express provides OTP (\texttt{/auth/request-login-otp}, \texttt{/auth/verify-login-otp}) and DB manager routes (\texttt{/api/tables}, inserts, reads).
\item Security: CORS middleware active; secrets via environment variables.
\end{itemize}

\textbf{Challenges:}
\begin{itemize}[leftmargin=*]
\item Firestore indexes for \texttt{arrayContains} queries on \texttt{username}/\texttt{phone}.
\item OTP delivery depends on SendGrid/Twilio configuration.
\item Aerospike scans need pagination for large datasets.
\end{itemize}

\section{Work Completed and Remaining Tasks}

\subsection{Completed Milestones}
\begin{itemize}[leftmargin=*]
\item Auth with role/status enforcement and step-up 2FA (OTP + trusted devices).
\item Firestore audit logging (admin/cashier) with device info collection.
\item Express backend with CORS, OTP endpoints, and DB manager across Firestore/Aerospike.
\item Aerospike integration: inserts and normalized scan reads.
\end{itemize}

\subsection{Outstanding Tasks and Next Steps}
\begin{itemize}[leftmargin=*]
\item Add pagination/filters to Aerospike read-all; document scan cost.
\item Add structured request logging and correlation IDs to backend.
\item Document Firestore security rules and indexes; parameterize backend URL in app config.
\item Implement server endpoint for password-reset OTP (client already wired).
\item Plan ML object detection integration (service boundary + data flow).
\end{itemize}

\section{Screenshot of the UI/UX}
\textbf{Image Placeholder: Primary Dashboard View}\\
Insert a high-quality screenshot of the main dashboard.\\
\emph{Figure 1: Navigation, key metrics display, and action buttons.}

\textbf{Image Placeholder: Data Input/Form View}\\
Insert a screenshot of a key form (login or OTP).\\
\emph{Figure 2: Fields, validation states, and feedback messages.}

\subsection{Screenshot 1: Primary Dashboard View}
\subsection{Screenshot 2: Data Input/Form View}

\section{Challenges and Mitigation}
\subsection{CORS in Browser Calls}
Resolved by global CORS middleware and OPTIONS handling in Express.

\subsection{OTP Delivery Reliability}
Integrated SendGrid/Twilio; generic responses avoid enumeration; monitor API results.

\subsection{Aerospike Set Enumeration/Namespace}
Limited to namespace \texttt{oipms}; validated connection and parsing of \texttt{aerospikeInfo('sets')}.

\subsection{RBAC Consistency}
Enforced cashier approval; role cached with safe invalidation.

\section{Conclusion and Next Reporting Period}
We achieved working RBAC, 2FA, auditing, and a dual-database backend. Next period: production hardening (logging, pagination, security docs) and ML module planning.

\end{document}
